---
title: "Project EDS-214"
author: "Hylaea Miller"
format: html
editor: visual
#| echo: false
#| warning: false
---

## Background

### Packages used on the project and function

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(patchwork)
library(here)
library(paletteer)

source(here::here("R/Function.R"))
```

## Data

The data are available through the Environmental Data Initiative. You can find them here: [(McDowell and International Institute Of Tropical Forestry (IITF) 2024)](https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-luq.20.4923064)

-   Quebrada one-Bisley (Q1) Chemistry Data
-   Quebrada two-Bisley (Q2) Chemistry Data
-   Quebrada three-Bisley (Q3) Chemistry Data
-   Puente Roto Mameyes (MPR) Chemistry Data

## Methods

### Reading, formatting names and creating columns for each site

```{r}

bq1 <- read_csv(here("data/QuebradaCuenca1-Bisley.csv")) %>% 
  clean_names() %>% 
  mutate(site = "BQ1")

bq2 <- read_csv(here("data/QuebradaCuenca2-Bisley.csv")) %>% 
  clean_names() %>% 
  mutate(site = "BQ2")


bq3 <- read_csv(here("data/QuebradaCuenca3-Bisley.csv")) %>% 
  clean_names() %>% 
  mutate(site = "BQ3")


prm <- read_csv(here("data/RioMameyesPuenteRoto.csv")) %>% 
  clean_names() %>% 
  mutate(site ="PRM")
```

### Combining the sites and cleaning the data

```{r}

comb_table <- bind_rows(bq1, bq2, bq3, prm) %>% 
  mutate(sample_date = lubridate::ymd(sample_date)) %>% 
  mutate(year = lubridate::year(sample_date)) %>% 
  select(sample_date, k, no3_n, mg, ca, nh4_n, site, year)  # selecting variables of interest
```

### Switching class of nutrients to numeric and filtering to years of interest

```{r}

comb_table$k <- as.numeric(comb_table$k)
comb_table$no3_n <- as.numeric(comb_table$no3_n)
comb_table$mg <- as.numeric(comb_table$mg)
comb_table$ca <- as.numeric(comb_table$ca)
comb_table$nh4_n <- as.numeric(comb_table$nh4_n)


comb_table_subset_date <- comb_table %>%
  filter(year >= '1988' & year <= '1994')
```

### Discovering the nine-week moving average of nutrient concentration

#### Applying the function to potassium concentration

```{r}
# Creating vectors to use in the function

dates <- comb_table_subset_date$sample_date

conc <- comb_table_subset_date$k

names(comb_table_subset_date)

# Creating a data set for the potassium concentration

df_k <- comb_table_subset_date %>%
  group_by(site) %>%
  mutate(conc_k = sapply(sample_date,
                         moving_average,
                         dates = sample_date,
                         conc = k,
                         win_size_wks = 9))
```

#### Applying the function to the nitrate-N concentration

```{r}
# Creating vectors to use in the function for the nitrate-N

dates <- comb_table_subset_date$sample_date

conc <- comb_table_subset_date$no3_n

names(comb_table_subset_date)

# Creating a data set for the nitrate-N concentration

df_no <- comb_table_subset_date %>%
  group_by(site) %>%
  mutate(conc_no = sapply(sample_date,
                          moving_average,
                          dates = sample_date,
                          conc = no3_n,
                          win_size_wks = 9))
```

#### Applying the function to the magnesium concentration

```{r}
# Creating vectors to use in the function for the magnesium

dates <- comb_table_subset_date$sample_date

conc <- comb_table_subset_date$mg



# Creating a data set for the magnesium concentration

df_mg <- comb_table_subset_date %>%
  group_by(site) %>%
  mutate(conc_mg = sapply(sample_date,
                          moving_average,
                          dates = sample_date,
                          conc = mg,
                          win_size_wks = 9))
```

#### Applying the function to the calcium concentration

```{r}
# Creating vectors to use in the function for the calcium 

dates <- comb_table_subset_date$sample_date

conc <- comb_table_subset_date$ca

# Creating a data set for the calcium concentration

df_ca <- comb_table_subset_date %>%
  group_by(site) %>%
  mutate(conc_ca = sapply(sample_date,
                          moving_average,
                          dates = sample_date,
                          conc = ca,
                          win_size_wks = 9))
```

#### Applying the function to the ammonia-N

```{r}
# Creating vectors to use in the function for the ammonia-N

dates <- comb_table_subset_date$sample_date

conc <- comb_table_subset_date$nh4_n

# Creating a data set for the  ammonium-N concentration

df_nh4_n <- comb_table_subset_date %>%
  group_by(site) %>%
  mutate(conc_nh4_n = sapply(sample_date,
                             moving_average,
                             dates = sample_date,
                             conc = nh4_n,
                             win_size_wks = 9))
```

## Results

### Plotting the moving average of the potassium concentration

```{r}

plot_k <- ggplot(df_k, aes(x = sample_date, y = conc_k, color = site)) +
  geom_line() +
  scale_y_continuous(limits = c(0.4, 1.6)) +
  scale_color_paletteer_d("wesanderson::Darjeeling2") +
  labs(title, x = "Year",
       y = "K mg l⁻¹") +
  theme_minimal()

print(plot_k)
```

### Plotting the moving average of the nitrate-N concentration

```{r}


plot_no <- ggplot(df_no, aes(x = sample_date, y = conc_no, color = site)) +
  geom_line() +
  scale_y_continuous(limits = c(5, 500)) +
  scale_color_paletteer_d("wesanderson::Darjeeling2") +
  labs(title, x = "Year",
       y = "NO₃-N ug l⁻¹") +
  theme_minimal()

print(plot_no)
```

### Plotting the moving average of the magnesium concentration

```{r}

plot_mg <- ggplot(df_mg, aes(x = sample_date, y = conc_mg, color = site)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 5)) +
  scale_color_paletteer_d("wesanderson::Darjeeling2") +
  labs(title, x = "Year",
       y = "Mg mg l⁻¹") +
  theme_minimal()

print(plot_mg)
```

### Plotting the moving average of the calcium concentration

```{r}
plot_ca <- ggplot(df_ca, aes(x = sample_date, y = conc_ca, color = site)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 10)) +
  scale_color_paletteer_d("wesanderson::Darjeeling2") +
  labs(title, x = "Year",
       y = "Ca mg l⁻¹") +
  theme_minimal()

print(plot_ca)

```

### Plotting the moving average of the ammonium-N concentration

```{r}

plot_nh4_n <- ggplot(df_nh4_n, aes(x = sample_date, y = conc_nh4_n, color = site)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 80)) +
  scale_color_paletteer_d("wesanderson::Darjeeling2") +
  labs(title, x = "Year",
       y = "NH4-N ug l⁻¹") +
  theme_minimal()

print(plot_nh4_n)
```

Plotting them together

```{r}
plot_k / plot_no / plot_mg / plot_ca / plot_nh4_n
```
